---
title: "Explore PL stat and variance partitioning data"
author: "KE Lotterhos"
date: "10/2/2020"
output:
  pdf_document: default
  html_document: default
---

## Update of Oct 5 post, with new data sent by Molly

setwd("/Users/lotterhos/Documents/GitHub/CnGV/notebook")
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
powdf <- read.csv("../results/Results_10.06.2020/Archive/Power_output_results.csv")
vardf <- read.csv("../results/Results_10.06.2020/Archive/Variance_output_results.csv")
PLdf <- read.csv("../results/Results_10.06.2020/Archive/PL_output_results.csv")
phendf <- read.csv("../results/Results_10.06.2020/Archive/phenotype_output_results.csv")
head(powdf)
head(vardf)
head(PLdf)
head(phendf)

dim(powdf)
dim(vardf)
dim(PLdf)


summary(powdf$row)
levels(as.factor(powdf$replicate))
summary(vardf$row)
summary(PLdf$row)

sum(!complete.cases(phendf))
dim(phendf)
```

# Explore PL stat

### subset to 2x2 cases
```{r}
cases <- powdf$row[powdf$n_env==2 & powdf$n_pop==2]
length(cases)

powdf2 <- powdf[powdf$row %in% cases,]
vardf2 <- vardf[vardf$row %in% cases,]

dim(powdf2)
head(PLdf)

newdf <- merge(powdf2, PLdf, by = "row")
dim(newdf)
str(newdf)

newdf$G1E1mean <- NA
newdf$G1E2mean <- NA
newdf$G2E1mean <- NA
newdf$G2E2mean <- NA
# to do: calc from phendf
i = 1001
cond <- which(phendf$row==i)
tapply(phendf$phen_corrected[cond],as.character(phendf$GE_factor[cond]), mean, na.rm=TRUE)
```

```{r}

## compare covariance and PL
  plot(newdf$covariance.x, newdf$PL)
  
  plot(newdf$covariance.x, newdf$PL, ylim=c(-1,2))
  abline(1,0)
  abline(0,0)
  abline(v=0)
  # this tells me that their claims that counter > 1 or <0 and cogradient between 0 and 1 is correct

# Compare covariance and GxE
  plot(newdf$GxE_emm.x, newdf$PL)
  plot(newdf$GxE_emm.x, newdf$PL, ylim=c(-1,2))
  
  plot(newdf$GxE_eta, newdf$PL, ylim=c(-1,2))
  summary(newdf$GxE_eta)
  # this tells me there PL metric has no relationship
  # with GxE
  
  plot(newdf$GxE_emm.x, newdf$GxE_omega)
  
# compare dela_E and new PL  
  plot(newdf$delta_E, newdf$PL, ylim=c(-1,2))
  # when delta_E is large, there is plasticity
  # when delta_E = 0, there is no plasticity or there is GxE
  # not very useful. It shows that PL is near 0 when delta_e is 0
  
# compare genetic differentiation and PL
  # denominator of PL minus the numerator is the average difference in phenotype between genotypes raised in a common environment (0.5((PAEA−PBEA)+(PAEB−PBEB))), and is therefore a measure of genetic differentiation.”
  plot(newdf$G1Emean - newdf$G2Emean, newdf$delta_E-newdf$delta_H)
  
    plot(abs(newdf$G1Emean - newdf$G2Emean), abs(newdf$delta_E-newdf$delta_H))
  
  plot(newdf$GxE_emm.x, newdf$delta_E-newdf$delta_H)
```

## Contrasting cases for PL
Show a case of countergradient variation with PL greater than 1, but plasticity induced response different sign as phenotypic divergence.
Show equivalent case of countegradient variation in opposite direction but with PL less than 0.
```{r}
b <- which(abs(round(newdf$PL)) > 100)
newdf[b,]
newdf[b,c("row","PL")]
```


Find a case with PL about 1 but huge GxE
```{r}
head(newdf)
a<- which(round(newdf$PL,1)==1.0 & newdf$GxE_omega > 0.6)
newdf[a,]
```

```{r}
i1 = 425
i1 <- 17238
i1 <- 16425
# from phendf, get phenotype values
cond1 <- which(newdf$row==i1)
newdf[cond1,]
cond <- which(phendf$row==i1)
a<- tapply(phendf$phen_corrected[cond],as.character(phendf$GE_factor[cond]), mean, na.rm=TRUE)
a

  plot(0:1, c(a[1:2]), ylim=c(-3,3), xaxt ="n", type="l", col="blue", xlab="", ylab="Standardized phenotype")
  mtext("E1", side=1, line=0, adj=0, col="blue")
  points(0:1, c(a[3:4]), ylim=c(-3,3), xaxt ="n", type="l", col="green", xlab="")
  mtext("E2", side=1, line=0, adj=1, col="green")

  text(0.3,a[1],"D_H", cex=0.5)
  arrows(0.4,a[1], 0.4, a[4], length=0.1)
  
  text(0.6, a[1], "D_E", cex=0.5)
  arrows(0.5, 0, 0.5, -newdf[cond1,"delta_E"], length=0.1)
```

## Explore variance partitioning
```{r}
head(powdf)
head(vardf)

# effect sizes
plot(powdf$covariance, 
     vardf$omega2[vardf$Variance_Component=="V_cov"],
     pch=19, 
     col=adjustcolor("blue",0.05))

plot(powdf$covariance, 
     vardf$omega2[vardf$Variance_Component=="V_g"],
     pch=19, 
     col=adjustcolor("blue",0.05))

plot(powdf$covariance, 
     vardf$omega2[vardf$Variance_Component=="V_e"],
     pch=19, 
     col=adjustcolor("blue",0.05))

plot(powdf$GxE_emm, 
     vardf$omega2[vardf$Variance_Component=="V_gxe"],
     pch=19, 
     col=adjustcolor("blue",0.05))

# P-values
plot(powdf$covariance_pvalue, 
     vardf$Pvalue[vardf$Variance_Component=="V_cov"],
     pch=19, 
     col=adjustcolor("blue",0.05))
    # p-values from the variance components are not to be trusted. Put on side burner for now.

plot(powdf$GxE_eta,  
     vardf$omega2[vardf$Variance_Component=="V_gxe"],
     pch=19, 
     col=adjustcolor("blue",0.1))
abline(0,1)
  # vardf$omega is actually eta

```

### Propose for the example figures where we show the raw data to add an insert showing the variance partitioning

omega^2 in variance partitioning code is actually eta^2 
```{r}
cond <- vardf$row==1001
vardf[cond,]
colr = c(grey(0.1), grey(.3), grey(.50), grey(.70), grey(.90))

barplot(vardf$omega2[cond],
        names= vardf$Variance_Component[cond],
        beside=FALSE,
        col=colr, las=2, xlab="", ylab="omega^2")

barplot(matrix(c(vardf$omega2[cond], rep(0,5)), ncol=2),
        xlim=c(0,2),
        beside=FALSE,
        col=colr, las=2, xlab="", ylab="omega^2")
legend("right",legend = rev(vardf$Variance_Component[cond]), fill=rev(colr))
```


