---
title: "New way of creating and standardizing phenotype data with variance partitioning"
author: "KE Lotterhos"
date: "8/18/2020"
output:
  pdf_document: default
  html_document: default
---

$V_P = V_G + V_E + V_{GE} + Cov_{GE}$

In a reciprocal transplant experiment, there are $g$ genotypes transplanted into $e$ environmental patches, for a total of $g*e = n_{GE}$ genotype-environment combinations. 

$g$ is levels of genotypes for $i = 1,2... g$
$e$ is levels of environments for $j = 1,2... e$

Assuming the equal sample sizes $r$ ($1, 2, ...k$) within each genotype-environment combination, could partitioning the variance be as simple as:

$V_G = re\sum_{i=1}^g (\bar{y_i} - \bar{y})^2$

$V_E = rg\sum_{j=1}^e (\bar{y_j} - \bar{y})^2$

$V_{GE} = r \sum_{i=1}^g \sum_{j=1}^e (\bar{y_{ij}} - \bar{y_i} - \bar{y_j} + \bar{y})^2$

$V_{Cov_{GE}}= (g-1)er\sum_{i=1}^g\sum_{j=1}^e(\bar{y_i} - \bar{y})(\bar{y_j} - \bar{y})I$

* I is an indicator variable that is 1 when the genotype $i$ originated from environment $j$ and 0 otherwise
* basically, each summation needs to count the same number of times to
be comparable as a SS on the same scale. If we have a 2x2 reciprocal transplant, we have 2 of 4 treatment combinations that count toward the sum (and 2 that should but do not). If we have a 4x4 reciprocal transplat, we have 4 of 16 reciprocal transplants that count toward the sum (and 12 that should but do not). We can get the SS to the right amount by multiplying (g-1)*e to get the same number of observations as the other SS calculations. At this point I'm only confident this will work in a reciprocal transplant case, would have to test g=4 and e=2 level cases etc.


$V_{error} = \sum_{i=1}^g \sum_{j=1}^e \sum_{k=1}^r (y_{ijk}-\bar{y}_{ij})^2$

Some evidence that this is behaving as expected:

if you set beta1 = 1 and beta2 = 1 and int = 0, you get equal amounts of variance explained for G, E, and CovGE for 2 pops, but CovGE increases with more pops

if you set beta1 = 1 and beta2 = 0.1 and int = 0, you get more variance explained for V_E, a little bit for V_G, and a decent amount for cov_GE

if you set beta1 = 0.1 and beta2 = 1 and int = 0, you get the same pattern as above but V_E and V_G switched

if you set beta1 or beta2 negative and the other positive, you get negative covGE

try adding noise and see if outputs make sense
* Adding noise can sometimes "create" some CovGE when there are enough pops and beta1=0 or beta2=0
* But for most part seems to work (?)

try increasing n_pop and see if outputs make sense
* increasing n_pop can lead to a greater amount of variance explained by CovGE and less by VG or VE




```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
  library("emmeans")
  library("lme4")
  library("rlist")
  library("dplyr")
```

# Step 1: create true phenotype data without error
```{r}

beta1 = 10 # the amount the phenotype changes across 1 value of the environment (i.e., the slope). This is essentially the amount/degree of phenotypic plasticity that is the same across genotypes.

n_genotypes = 10 # the number of genotypes collected from different locations
n = 10 # sample size
beta2 = 10 # the amount the phenotype changes from one genotype to the next. This is essitially the increase intercept from one genotype to the next.

scale = 0.01 # the scale for sd_noise within pops compared to sd_means among pops
# I'll let you explore the space, but I think the upper would be scale = 1
# I'd start with scale = 1 and scale = 1/2 and see what that looks like

sd_int = 0#n_genotypes # sd of the interaction terms that will be drawn
    #  beta1 <- 1
    #  beta2 <- 1
    
    # Generate data
    
    n_environments = n_genotypes # the number of common garden environments that the genotypes are grown in. Basically, genotype[i] orginates from environment[i]. This matching between genotypes and environment is important for the the cov(G,E) calculation. Here, we only consider the case of the complete reciprocal transplant experiment.

    ## Create levels of genotypes and environments ####
    G = rep(1:n_genotypes, each=n*n_genotypes) # '0' n times, '1' n times
    E = rep(1:n_environments,times=n*n_genotypes) # '0'x50, '1'x50, '0'x50, '1'x50
    #set.seed(97)
    

    ## Create interaction effect for each level of both factors ####
    # In this case we assume the GxE interactions are a 
    # normally distributed random variable with a mean of 0
    # As the sd increases, so does the GxE among treatments
    if(sd_int == 0){
      int = 0
    }else{
    int <- rnorm(n_genotypes * n_environments, 0, sd=sd_int)
    }
    # no GxE
 
     # this sd determines the amount of GxE
    int_df <- data.frame(expand.grid(G=1:n_genotypes, E=1:n_environments), 
                         int)
    
    ### Create the model dataframe ####
    model_df <- data.frame(G, E)
    model_df$GE_factor <- paste0("G",model_df$G, "E", model_df$E)
    model_df <- merge(model_df, int_df)
    model_df <- model_df[order(model_df$G, model_df$E),]
    dim(model_df)
    head(model_df, 30)
    
    # Generate phenotype data using the regression equation ####
    model_df$GE_true = beta1*model_df$E + beta2*model_df$G + model_df$int
    
    G_true <- data.frame(G=1:n_genotypes, G_true=tapply(model_df$GE_true, model_df$G, mean))
    E_true <- data.frame(E=1:n_genotypes , E_true=tapply(model_df$GE_true, model_df$E, mean))
    
    model_df1 <- merge(model_df,G_true)
    model_df2 <- merge(model_df1,E_true)
    model_df <- model_df2
    
    model_df$mean_true <- mean(model_df$GE_true)
    
    model_df$int_true <- model_df$mean_true + model_df$GE_true - 
                        model_df$G_true -   model_df$E_true
    head(model_df)
    tail(model_df)
    par(mfrow=c(1,1))
    plot(model_df$int, model_df$int_true)
    abline(0,1)
```

Note that the way we add "int" to create phenotypic variation is
not the same way that the actual interaction is calculated from the data. Our simulations create an interaction, but as we create an interaction we also change the G-means and E-means, and so the calculated interaction is a bit different.

# Step 2: standardize true phenotype data
```{r}
GE_true_means <- tapply(model_df$GE_true, model_df$GE_factor, mean)

model_df$GE_stn_true <- (model_df$GE_true - mean(GE_true_means))/sd(GE_true_means)

GE_stn_true_means <- tapply(model_df$GE_stn_true, model_df$GE_factor, mean)

    G_stn_true <- data.frame(G=1:n_genotypes, G_stn_true=tapply(model_df$GE_stn_true, model_df$G, mean))
    E_stn_true <- data.frame(E=1:n_genotypes , E_stn_true=tapply(model_df$GE_stn_true, model_df$E, mean))
    
    model_df1 <- merge(model_df,G_stn_true)
    model_df2 <- merge(model_df1,E_stn_true)
    model_df <- model_df2

    model_df$mean_stn_true <- mean(model_df$GE_stn_true)
    
    model_df$int_stn_true <- model_df$mean_stn_true + model_df$GE_stn_true - 
                        model_df$G_stn_true -   model_df$E_stn_true
    head(model_df)
    tail(model_df)
    par(mfrow=c(1,1))
    plot(model_df$int_true, model_df$int_stn_true)
      # awesome
```


# Step 3: add error to stnd phenotype data 
scaled to the variation in the phenotype means
```{r}
sd(GE_stn_true_means)
# I guess this will always be 1 because we standardize.
sd_noise <- sd(as.numeric(GE_stn_true_means))*scale
model_df$e = rnorm(n*n_genotypes*n_environments, 0, sd=sd_noise) # Random noise

model_df$phenotype <- model_df$GE_stn_true + model_df$e
head(model_df)
```

# Step 4: plot the pattern so we can see what it looks like
```{r}   
   par(mfrow=c(1,1), mar=c(4,4,1,1))
    plot(model_df$phenotype~model_df$E, col=0, xlim=c(1,n_genotypes+1), ylim=c(-3,3))
    for (m in 1:n_genotypes){
      x <- model_df$E[model_df$G==m]+0.05*(m-1)
      points(model_df$phenotype[model_df$G==m]~x, col=m+1)
      p_means <- tapply(model_df$phenotype[model_df$G==m],model_df$E[model_df$G==m], mean)
      points(p_means~as.numeric(names(p_means)),
             col=m+1, type="l")
    }
```


# Step 5: pretend we start with noisy simulated data
Go through the steps we will go through in the study
```{r}
# Step A: standardize by mean(GE_means) and sd(GE_means)
phen_GE_obs_means <- tapply(model_df$phenotype, model_df$GE_factor, mean)

plot(GE_stn_true_means, phen_GE_obs_means) #sanity check
abline(0,1)

model_df$phenotype_stn <- (model_df$phenotype-mean(phen_GE_obs_means))/sd(phen_GE_obs_means)

plot(model_df$phenotype, model_df$phenotype_stn)
abline(0,1)

# Step B: calculate observed G_means and E_means and interaction
    G_stn_est <- data.frame(G=1:n_genotypes, G_stn_est=tapply(model_df$phenotype_stn, model_df$G, mean))
    E_stn_est <- data.frame(E=1:n_genotypes , E_stn_est=tapply(model_df$phenotype_stn, model_df$E, mean))
    
    GE_stn_est_means <- tapply(model_df$phenotype_stn, model_df$GE_factor, mean)
    GE_stn_est_means_df <- data.frame(GE_factor=names(GE_stn_est_means) , GE_stn_est=GE_stn_est_means)
    
    model_df1 <- merge(model_df,G_stn_est)
    model_df2 <- merge(model_df1,E_stn_est)
    model_df3 <- merge(model_df2, GE_stn_est_means_df)
    model_df <- model_df3
    
    plot(model_df$GE_stn_true, model_df$GE_stn_est)
    abline(0,1) # looks good
    
    model_df$mean_stn_est <- mean(model_df$phenotype_stn)
    
    # Calculate interaction
    model_df$int_stn_est <- model_df$mean_stn_est + model_df$GE_stn_est - 
                        model_df$G_stn_est -   model_df$E_stn_est
```

# Step 6: Compare true values to estimated values
```{r}
head(model_df)
(true_int <- mean(abs(model_df$int_stn_true)))

(obs_int <- mean(abs(model_df$int_stn_est)))
```


# New step 6.5: make ANOVA table with CovGE

Here, I didn't write loops to calculate the equations, so the code will look different from the equations.

TO DO: check to code the loops as written in the equations and make sure it gives the same answer as what I calculated here

```{r}
head(model_df)

V_G_SS = sum((model_df$G_stn_est-model_df$mean_stn_est)^2)
V_E_SS = sum((model_df$E_stn_est-model_df$mean_stn_est)^2)
V_GE_SS = sum(model_df$int_stn_est^2)
V_error = sum((model_df$phenotype_stn - model_df$GE_stn_est)^2)

model_df$I = model_df$E==model_df$G

# Covariance pattern (ignore the 0,0 points)
par(mfrow=c(2,1))
plot(model_df$G_stn_est*model_df$I, model_df$E_stn_est*model_df$I)
plot(model_df$G_stn_est*model_df$I, model_df$E_stn_est*model_df$I, xlim=c(-1,1), ylim=c(-1,1))

V_Cov_GE <-  (nlevels(factor(model_df$G))-1)*
  nlevels(factor(model_df$E))*
  sum((model_df$G_stn_est-model_df$mean_stn_est)*(model_df$E_stn_est-model_df$mean_stn_est)*model_df$I)

SS <- round(rbind(V_G_SS, V_E_SS, V_GE_SS, V_Cov_GE, V_error),2)
omega2 <- round(abs(SS)/sum(abs(SS)),2)
data.frame(SS, abs(SS), omega2)
```


# Step 7: pretend we start with noisy empirical data
Let's back transform the noisy data, then go through the steps and see what happens
```{r}
# what we did to standardize: model_df$GE_stn_true <- (model_df$GE_true - mean(GE_true_means))/sd(GE_true_means)
model_df2 <- model_df[,1:3]
model_df2$phenotype2 <- model_df$phenotype*sd(GE_true_means)+mean(GE_true_means) #backtransform


# Step A: standardize by mean(GE_means) and sd(GE_means)
phen_GE_obs_means <- tapply(model_df2$phenotype2, model_df2$GE_factor, mean)

model_df2$phenotype2_stn <- (model_df2$phenotype2-mean(phen_GE_obs_means))/sd(phen_GE_obs_means)

plot(model_df2$phenotype2, model_df2$phenotype2_stn) # straight line check

# Step B: calculate observed G_means and E_means and interaction
    G_stn_est <- data.frame(G=1:n_genotypes, G_stn_est=tapply(model_df2$phenotype2_stn, model_df2$G, mean))
    E_stn_est <- data.frame(E=1:n_genotypes , E_stn_est=tapply(model_df2$phenotype2_stn, model_df2$E, mean))
    
    GE_stn_est_means <- tapply(model_df2$phenotype2_stn, model_df2$GE_factor, mean)
    GE_stn_est_means_df <- data.frame(GE_factor=names(GE_stn_est_means) , GE_stn_est=GE_stn_est_means)
    
    model_dfa <- merge(model_df2,G_stn_est)
    model_dfb <- merge(model_dfa,E_stn_est)
    model_dfc <- merge(model_dfb, GE_stn_est_means_df)
    head(model_dfc)
    model_df2 <- model_dfc
    
    plot(model_df2$phenotype2_stn, model_df2$GE_stn_est)
    abline(0,1) # looks good
    
    model_df2$mean_stn_est <- mean(model_df2$phenotype2_stn)
    
    # Calculate interaction
    model_df2$int_stn_est <- model_df2$mean_stn_est + model_df2$GE_stn_est - 
                        model_df2$G_stn_est -   model_df2$E_stn_est
    
    
```

# Step 8: Compare true values to estimated values
```{r}
head(model_df)
(true_int <- mean(abs(model_df$int_stn_true)))
# Although we call this "true_int", when int=0 this should be 0
# note this is a shortcut that only works with equal sample sizes

(obs_int_sim <- mean(abs(model_df$int_stn_est)))
# when int=0, this will increase as the within-population mean gets less accurate

(obs_int_emp <- mean(abs(model_df2$int_stn_est)))
```

# Drop the mic.