---
title: "Multiple Populations for CnGv and GxE"
author: "KE Lotterhos"
date: "1/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The purpose of this code is to provide a more general framework for modeling phenotypes in a reciprocal transplant experiment with more than two locations, when there is both GxE and cov(G,E) affecting the phenotype.

Basically, we model the phenotype as an additive function of:
- an intercept that is based on the location/environment the phenotype was collected from. Intercepts increase or decrease in a linear manner across the environment.
- a slope that determines the plastic response of the phenotype across the environment
- an interaction term that determines how the phenotype of genotype $i$ in environment $j$ deviates from the main effects
- random error or noise

```{r cars}
beta1 = 1 # the amount the phenotype changes across 1 value of the environment (i.e., the slope). This is essentially the amount/degree of phenotypic plasticity that is the same across genotypes.

beta2 = -1 # the amount the phenotype changes from one genotype to the next. This is essitially the increase intercept from one genotype to the next.

(cov_GE_approx = beta1 * beta2) 
  # will be correlated with the magnitude of cov(G,E)

# Generate data
n = 15 # sample size
n_genotypes = 10 # the number of genotypes collected from different locations
n_environments = n_genotypes # the number of common garden environments that the genotypes are grown in. Basically, genotype[i] orginates from environment[i]. This matching between genotypes and environment is important for the the cov(G,E) calculation. Here, we only consider the case of the complete reciprocal transplant experiment.

## Create levels of genotypes and environments ####
G = rep(1:n_genotypes, each=n) # '0' n times, '1' n times
E = rep(1:n_environments,times=n) # '0'x50, '1'x50, '0'x50, '1'x50
set.seed(97)
e = rnorm(n*n_genotypes, 0, sd=0.01) # Random noise, with standard deviation of 1

## Create interaction effect for each level of both factors ####
# In this case we assume the GxE interactions are a 
# normally distributed random variable with a mean of 0
# As the sd increases, so does the GxE among treatments
int <- rnorm(n_genotypes * n_environments, 0, sd=1)
  # this sd determines the amount of GxE
int_df <- data.frame(expand.grid(G=1:n_genotypes, E=1:n_environments), 
                     int)

### Create the model dataframe ####
model_df <- data.frame(G, E, e)
model_df <- merge(model_df, int_df)

# Generate phenotype data using the regression equation ####
phenotype = beta1*model_df$E + beta2*model_df$G  + model_df$e + model_df$int
plot(phenotype~model_df$E, col=model_df$G+1)
```

