---
title: "How to calculate the mean"
author: "KE Lotterhos"
date: "3/9/2021"
output:
  pdf_document: default
  html_document: default
---

```{r}

# Different ways of calculating means ####
getmeans <- function(ind=1:length(a), true){
  obs_all <- mean(b[ind]) # This may or may not be what your eMM does
  obs_mean_GE <- mean(tapply(b[ind],a[ind], mean))
  obs_mean_G <- mean(tapply(b[ind],G_lev[ind], mean))
  obs_mean_E <- mean(tapply(b[ind],E_lev[ind], mean))
  obs_mean_cGE <- mean(c(obs_mean_G, obs_mean_E))  # Note they count unequally in the common garden N x 2 case
  out <- cbind(obs_all, obs_mean_GE, obs_mean_G,obs_mean_E, obs_mean_cGE)
  round(out - true, 4)
}
```

# consider 2x2 case ####

First, consider the 2x2 case, with different GE means for each level. Compare results for imbalance in the level with the lowest mean, and imbalance in the level with the highest mean. 

```{r}
a <- rep(c("G1E1","G1E2", "G2E1", "G2E2"), each=20)
G_lev <- substr(a, start=1, stop=2)
E_lev <- substr(a, start=3, stop=4)
b <- c(rnorm(length(a)/4, -1), rnorm(length(a)/4, 0), rnorm(length(a)/4, 1), rnorm(length(a)/4, 2))
hist(b)
boxplot(b~a)

true1 <- mean(c(-1,0,1,2)) # true mean

getmeans(1:length(a), true1) # no imbalance
getmeans(15:length(a), true1) # imbalance in first category
getmeans(1:(length(a)-15), true1) # imbalance in last category

par(mar=c(10,3,1,1))
bob <- replicate(10000,getmeans(c(sample(1:20,5),20:length(a)), true1),simplify = TRUE)
rownames(bob) <- c("obs_all", "obs_mean_GE", "obs_mean_G", "obs_mean_E", "obs_mean_cGE")
#head(bob)
boxplot(t(bob), las=2)
abline(h=0)

bob <- replicate(1000,getmeans(c(1:(length(a)-20), sample((length(a)-20):length(a),5)), true1),simplify = TRUE)
#head(bob)
rownames(bob) <- c("obs_all", "obs_mean_GE", "obs_mean_G", "obs_mean_E", "obs_mean_cGE")
#head(bob)
boxplot(t(bob), las=2)
abline(h=0)
```

# consider 2x4 common garden case ####
First, consider the 2x4 case, with different GE means for each level. Compare results for imbalance in the level with the lowest mean, and imbalance in the level with the highest mean. 
```{r}
a <- rep(c("G1E1","G1E2", "G2E1", "G2E2", "G3E1", "G3E2", "G4E1", "G4E2"), each=20)
G_lev <- substr(a, start=1, stop=2)
E_lev <- substr(a, start=3, stop=4)
b <- c(rnorm(length(a)/8, -1), 
       rnorm(length(a)/8, 0), 
       rnorm(length(a)/8, 1), 
       rnorm(length(a)/8, 2), 
       rnorm(length(a)/8, -1), 
       rnorm(length(a)/8, 0), 
       rnorm(length(a)/8, 1), 
       rnorm(length(a)/8, 2))
hist(b)
boxplot(b~a)

par(mar=c(10,3,1,1))

true2 <- mean(c(-1,0,1,2, -1, 0, 1,2,-1,0,1,2)) # true mean

bob <- replicate(1000,getmeans(c(sample(1:20,5),20:length(a)), true2),
                 simplify = TRUE)
rownames(bob) <- c("obs_all", "obs_mean_GE", "obs_mean_G", "obs_mean_E", "obs_mean_cGE")
#head(bob)
boxplot(t(bob), las=2)
abline(h=0)

bob <- replicate(1000,getmeans(c(1:(length(a)-20), 
                                sample((length(a)-20):length(a),5)), true2),
                 simplify = TRUE)
rownames(bob) <- c("obs_all", "obs_mean_GE", "obs_mean_G", "obs_mean_E", "obs_mean_cGE")
#head(bob)
boxplot(t(bob), las=2)
abline(h=0)
```
