---
title: "co-counter gradient variation"
author: "Katie Lotterhos, Molly Albecker"
date: "5/30/2019"
output: html_document
---
```{r}
library("parallel")
library("doParallel")
library("tidyverse")
```

### Simulations ###

```{r}

## Establish dataframes
input <- list(
  "type" = "cogv",#c("cogv","cngv"),
  "intercept_G1" = 0,
  "slope_G1" = 0.5,#seq(from = -1, to = 1, by = 0.1),
  "intercept_G2" = 1.5,#seq(from = -5, to = 5, by = 2),
  "slope_G2" = 0.75,#seq(from = -1, to = 1, by = 0.1),
  "sd" = 1,#seq(from = 0, to = 1, by = 0.5),
  "reps" = 3,#seq(from = 5, to = 10, by = 2),
  "env_num" = 5,#c(2,5,10), 
  "env" = NA,
  "G1_env" = NA,
  "G2_env" = NA,
  "true_covGE" = NA, 
  "is.GxE" = NA, 
  "slope_diff" = NA) 


output_df = data.frame()
```

```{r Lord of the Function}


## One function to rule them all
the_ring <- function(input_df){
  
  # Starting Parameters
  type = input_df$type
  intercept_G1 = input_df$intercept_G1
  slope_G1 = input_df$slope_G1
  intercept_G2 = input_df$intercept_G2
  slope_G2 = input_df$slope_G2
  sd = input_df$sd
  reps = input_df$reps
  G1_env = input_df$G1_env
  G2_env = input_df$G2_env
  env_num = input_df$env_num
  env = input_df$env
  true_covGE = input_df$true_covGE
  is.GxE = input_df$is.GxE
  slope_diff = input_df$slope_diff
  
  for(a in 1:env_num){
    for(b in 1:length(unique(input_df$type))){
    
    env_num_temp = env_num[a]
    type_temp = type[b]
  
    # Set environmental parameters
    if(env_num_temp == 2){env = c(1,5)}else if (env_num_temp == 5){env = c(1,2,3,4,5)}else{env = seq(from = 0.5, to = 5, by = 0.5)}
    
    # Match genotypes to phenotypes
    if(type_temp == "cogv"){((G1_env= 1) & (G2_env = 5)) }else{ ((G1_env= 5) & (G2_env = 1))}
  
  
    for(c in 1:length(env)){
      for(d in 1:length(slope_G1)){
        for(e in 1:length(intercept_G2)){
          for(f in 1:length(slope_G2)){
            for(g in 1:length(sd)){
              for(h in 1:length(reps)){
                  
                  # GxE logical
                  if(slope_G1[d] == slope_G2[f]){is.GxE = "no"} else {is.GxE = "yes"}
                  
                  # Difference in slope
                  slope_diff = (slope_G2[f] - slope_G1[d])
                  
                  dat_temp = data.frame()
                  
                  # Generate data 
                  phen1 = c(replicate(reps[h], (intercept_G1 + slope_G1[d] * env + (rnorm(env_num_temp, 0, sd[g])))))
                  temp1 = data.frame("phen" = phen1,
                                     "gen" = rep("G1", length(env_num_temp*reps[h])),
                                     "env" = rep(env, reps[h]),
                                     "E_mean" = rep(G1_env,length(env_num_temp*reps[h])),
                                     "sample_size" = rep(reps[h], length(env_num_temp)*reps[h]))
                    
                  phen2 = c(replicate(reps[h], (intercept_G2[e] + slope_G2[f] * env + (rnorm(env_num_temp, 0, sd[g])))))
                  temp2 = data.frame("phen" = phen2,
                                     "gen" = rep("G2", length(env_num_temp*reps[h])),
                                     "env" = rep(env, reps[h]),
                                     "E_mean" = rep(G2_env,length(env_num_temp*reps[h])),
                                     "sample_size" = rep(reps[h], length(env_num_temp*reps[h])))
                  dat_temp = rbind(temp1, temp2)
                  
                  # Standardize generated data
                  dat_avg = mean(dat_temp$phen) 
                  dat_std = sd(dat_temp$phen)
                  dat_temp$phen_corrected = ((dat_temp$phen-dat_avg)/dat_std)
                  
                  # Model and outputs
                  
                  if(is.GxE == "no"){
                    test_temp = lm(phen_corrected ~ env + gen, data = dat_temp)
                    lm_type = "No_GxE"
                    slope_G1_lm = summary(test_temp)[[4]][2,1]
                    slope_G2_lm = summary(test_temp)[[4]][2,1]
                    intercept_G1_lm = summary(test_temp)[[4]][1,1]
                    intercept_G2_lm = (summary(test_temp)[[4]][1,1]+summary(test_temp)[[4]][3,1])
                    slope_GxE_lm = 0
                    lowCI_G1_int_lm = confint(test_temp)[1]
                    highCI_G1_int_lm = confint(test_temp)[4]
                    lowCI_G2_int_lm = confint(test_temp)[3]
                    highCI_G2_int_lm = confint(test_temp)[6]
                    lowCI_G1_slope_lm = confint(test_temp)[2]
                    highCI_G1_slope_lm = confint(test_temp)[5]
                    lowCI_G2_slope_lm = confint(test_temp)[2]
                    highCI_G2_slope_lm = confint(test_temp)[5]
                    E_R2 = summary(aov(test_temp))[[1]][1,2]/sum(summary(aov(test_temp))[[1]][,2])
                    G_R2 = summary(aov(test_temp))[[1]][2,2]/sum(summary(aov(test_temp))[[1]][,2])
                    GxE_R2 = 0
                    }else{
                    test_temp = lm(phen_corrected ~ env * gen, data = dat_temp)
                    lm_type = "GxE"
                    slope_G1_lm = summary(test_temp)[[4]][2,1]
                    slope_G2_lm = (summary(test_temp)[[4]][2,1]+summary(test_temp)[[4]][4,1])
                    intercept_G1_lm = summary(test_temp)[[4]][1,1]
                    intercept_G2_lm = (summary(test_temp)[[4]][1,1]+summary(test_temp)[[4]][3,1])
                    slope_GxE_lm = summary(test_temp)[[4]][4,1]
                    lowCI_G1_int_lm = confint(test_temp)[1]
                    highCI_G1_int_lm = confint(test_temp)[5]
                    lowCI_G2_int_lm = confint(test_temp)[3]
                    highCI_G2_int_lm = confint(test_temp)[7]
                    lowCI_G1_slope_lm = confint(test_temp)[2]
                    highCI_G1_slope_lm = confint(test_temp)[6]
                    lowCI_G2_slope_lm = confint(test_temp)[4]
                    highCI_G2_slope_lm = confint(test_temp)[8]
                    E_R2 = summary(aov(test_temp))[[1]][1,2]/sum(summary(aov(test_temp))[[1]][,2])
                    G_R2 = summary(aov(test_temp))[[1]][2,2]/sum(summary(aov(test_temp))[[1]][,2])
                    GxE_R2 = summary.aov(test_temp)[[1]][3,2]/sum(summary.aov(test_temp)[[1]][,2])
                    }
                  
                  # To Estimate Covariance
                  cov_temp <- expand.grid(gen = unique(dat_temp$gen),
                                          env = seq(from = as.integer(min(dat_temp$env)), to = as.integer(max(dat_temp$env),by=1)),
                                          phen_corrected = NA)
                  mm = model.matrix(terms(test_temp),cov_temp)
                  cov_temp$phen_corrected = mm %*% coef(test_temp) # Predicted phenotypes
                  pvar1 <- diag(mm %*% tcrossprod(vcov(test_temp),mm))
                  cov_temp <- data.frame(cov_temp,
                                         plo = cov_temp$phen_corrected-1.96*sqrt(pvar1),
                                         phi = cov_temp$phen_corrected+1.96*sqrt(pvar1))
                  
                  # G_means
                  E_hat = mean(env)
                  G_means = filter(cov_temp, env == E_hat)
                  
                  # E_means
                  E_mean_temp = data.frame()
                  for(j in 1:length(unique(dat_temp$E_mean))){
                    E = unique(dat_temp$E_mean)[j]
                    E_temp = cov_temp[which(cov_temp$env == E),]
                    E_mean = mean(E_temp$phen_corrected)
                    E_mean_lowCI = mean(E_temp$plo)
                    E_mean_highCI = mean(E_temp$phi)
                    E_mean_temp1 = data.frame("E_mean" = E_mean,
                                              "E_mean_lowCI" = E_mean_lowCI,
                                              "E_mean_highCI" = E_mean_highCI,
                                              "env" = unique(E_temp$env))
                    E_mean_temp = rbind(E_mean_temp, E_mean_temp1)
                    }
                  
                  # Data output
                  RECODE TO USE ORIGINAL DATA - DONT FUDGE IT
                  CONF INT SHOULD COME FROM BOOTSTRAP - USE SAME FUNCTION ON THIS DATASET FOR EACH LOOP
                  AS WE USE ON REAL DATASET
                  TAKE DATASET YOU CREATE FROM THE TOP AND USE BOOTSTRAP
                  PVALUE FOR COV WILL COME FROM PERMUTATION
                  Cov_matrix = data.frame("gen" = unique(dat_temp$gen),
                                          "env" = as.factor(unique(dat_temp$E_mean)),
                                          "G_means" = G_means$phen_corrected,
                                          "G_means_lowCI" = G_means$plo,
                                          "G_means_highCI" = G_means$phi,
                                          "E_means" = E_mean_temp[,1],
                                          "E_mean_lowCI" = E_mean_temp[,2],
                                          "E_mean_highCI" = E_mean_temp[,3])
                  
                  if(type_temp == "cogv"){
                  est_covGE_lm = cov(Cov_matrix$G_means,Cov_matrix$E_means)
                  }else{
                  est_covGE_lm = cov(Cov_matrix$G_means,-(Cov_matrix$E_means))
                  }
                  
                  output_df. <- data.frame("lm_type" = lm_type,
                                           "type" = type_temp,
                                           "sd" = sd[g],
                                           "slope_G1" = slope_G1[e],
                                           "slope_G2" = slope_G2[f],
                                           "intercept_G1" = intercept_G1,
                                           "intercept_G2" = intercept_G2[e],
                                           "env_num" = env_num_temp,
                                           "sample_size" = reps[h],
                                           "est_covGE_lm" = est_covGE_lm,
                                           "GxE_R2" = GxE_R2,
                                           "slope_G1_lm" = slope_G1_lm,
                                           "slope_G2_lm" = slope_G2_lm,
                                           "intercept_G1_actual" = intercept_G1,
                                           "intercept_G1_lm" = intercept_G1_lm,
                                           "intercept_G2_actual" = intercept_G2[e],
                                           "intercept_G2_lm" = intercept_G2_lm, 
                                           "slope_GxE_lm" = slope_GxE_lm,
                                           "lowCI_G1_int_lm" = lowCI_G1_int_lm,
                                           "highCI_G1_int_lm" = highCI_G1_int_lm,
                                           "highCI_G2_int_lm" = highCI_G2_int_lm,
                                           "lowCI_G1_slope_lm" = lowCI_G1_slope_lm,
                                           "highCI_G1_slope_lm" = highCI_G1_slope_lm,
                                           "lowCI_G2_slope_lm" = lowCI_G2_slope_lm,
                                           "highCI_G2_slope_lm" = highCI_G2_slope_lm)
                  LOW_CI_COVGE
                  HIGH_CI_COVGE
                  P_COVGE (check if right tail p = 1-p)
                  output_df = rbind(output_df,output_df.)
              }
            }
          }
        }
      }
    }
    }
  }
   return(output_df)
  }

test_dat = the_ring(input)
ggplot(data = dat_temp, aes(x = env, y = phen_corrected, group = gen)) + 
  geom_smooth() + theme_classic() + geom_hline(yintercept = 0.7,lty = 2) + 
  geom_hline(yintercept = 0.92,col = 2) + geom_hline(yintercept = -0.92,col = 2) +
  geom_hline(yintercept = -0.7,lty = 2) + geom_vline(xintercept = 3,lty=2)+
  geom_vline(xintercept = 1,col=2)+geom_vline(xintercept = 5,col=2)

test_dat[2000:2050,]
```

```{r Simple linear model}
data_generation_noGxE <- function(base_intercept, int_eff, b1, env, sd, replicates){
    
  lmdat = data.frame()
  
  for(p in 1:length(reps)){
    for(i in 1:length(int_eff)){
      for(j in 1:length(sd)){
      
      # Counter
      cat(p,i,j, "\n") 

      # Establish parameters for models
      eff_increase = int_eff[i]
      sd_temp = sd[j]
      rp = reps[p]
      
      # Generate data
      phen_base = c(replicate(rp, (base_intercept + b1 * env + (rnorm(length(env), 0, sd_temp)))))
      base_temp = data.frame("phen" = phen_base,
                             "gen" = rep("G1", length(env*rp)),
                             "env" = rep(env, length(rp)),
                             "rep" = rep(rp, length(env*rp)))
                      
      phen2 = c(replicate(rp, ((base_intercept+eff_increase) + b1 * env + (rnorm(length(env), 0, sd_temp)))))
      change_temp = data.frame("phen" = phen2,
                             "gen" = rep("G2", length(env*rp)),
                             "env" = rep(env, length(rp)),
                             "rep" = rep(rp, length(env*rp)))
      dat_temp = rbind(base_temp, change_temp)
      
      # Model
      test_temp = lm(phen ~ env * gen, data = dat_temp)
      pval = summary(test_temp)$coefficients[,4][3]
      lwr_ci = confint(test_temp)[3]
      upr_ci = confint(test_temp)[6]

      lmdat. = data.frame("std.dev" = sd_temp,
                          "reps" = rp,
                          "effect.size" = eff_increase,
                          "p.val" = pval,
                          "lwr.ci" = lwr_ci,
                          "upr.ci" = upr_ci)
      lmdat = rbind(lmdat,lmdat.)
      }
    }
  }
  return(data.frame(lmdat))
}
```

```{r}
simnum = c(1:5)
simdat <- replicate(length(simnum), data_generation_noGxE(base_intercept, int_eff, b1, env, sd, reps), simplify=TRUE) 

# Convert array into dataframe
dat_combined = data.frame()

for(i in 1:length(simnum)){
  val = simnum[i]
  tempdat = as.data.frame(simdat[,i])
  tempdat = data.frame("iteration" = rep(val,nrow(tempdat)),
                       tempdat[,c(1:6)])
  dat_combined = rbind(dat_combined,tempdat)
  }
  
dat_combined$overunder = 0
for(j in 1:nrow(dat_combined)){
  if (dat_combined$lwr.ci[j] < 0) {dat_combined$overunder[j] = 0} else {dat_combined$overunder[j] = 1}
  }
newdat = dat_combined
```

```{r GetPower Function}

## Function to calculate power based on output 
detach(package:plyr)

powerdat = data.frame()
  for(i in 1:length(unique(newdat$reps))){
    rep = unique(newdat$reps)[i]
    repdat = filter(newdat, reps == rep)
    for(j in 1:length(unique(repdat$std.dev))){
      sdev = unique(newdat$std.dev)[j]
      repdat2 = filter(repdat, std.dev == sdev)
      for(k in 1:length(unique(repdat2$effect.size))){
        effsize = unique(repdat2$effect.size)[k]
        repdat3 = filter(repdat2, effect.size == effsize)
        
        powerdat. = data.frame("reps" = unique(repdat3$reps),
                                "std.dev" = unique(repdat3$std.dev),
                                "effect.size" = unique(repdat3$effect.size),
                                "numer" = sum(repdat3$overunder),
                                "denom" = nrow(repdat3))
        powerdat = rbind(powerdat,powerdat.)
      }
    }
  }

powerdat$power = (powerdat$numer/powerdat$denom) 
```
*Power Heatmap*

Power is the proportion of replicates whose 95% CI do not include 0

```{r}

## Plot Heatmap

simplot = ggplot(data = powerdat, aes(x = reps, y = effect.size))+ #reorder(std.dev,-std.dev)
  geom_tile(aes(fill = power), colour = "white") + scale_fill_gradient(low = "white", high = "steelblue")+
  theme_classic()+ 
  #ylab("Difference in G and E") + xlab("Sample Size")+
  theme_bw(base_size = 30, base_family = "Helvetica")+ 
  theme(strip.background =element_rect(fill="white"))+
  theme(strip.text = element_text(colour = 'black')) +
  theme(plot.background = element_blank(),panel.grid.major = element_blank(),panel.grid.minor = element_blank())+
  theme(axis.ticks = element_line(colour = "black"))+
  theme(axis.text= element_text(colour = "black"))+
  labs(colour = "Sample Size")+
  theme(legend.position="bottom")+
  theme(legend.title = element_text(size = 14),
          legend.text = element_text(size = 14))
  #theme(legend.position="none")
simplot

require("scatterplot3d")
require("rgl")

scatterplot3d(x = powerdat$reps, y= powerdat$power, z = powerdat$effect.size)
plot3d(powerdat$std.dev,powerdat$power,powerdat$effect.size)
wireframe(powerdat$power~newdat0$effect.size*powerdat$std.dev,colorkey=TRUE,drape=TRUE,
          scales=list(arrows=FALSE), zlim=c(0,1))
```

Plots for heuristics

```{r}

# Cogradient, No GxE

b0_1 = 1 #Intercept Group 1
b0_2 = 3 #Intercept Group 2
b1_1 = 0.1 #Slope Group 1
b1_2 = 0.1 #Slope Group 2

phen1 = b0_1 + b1_1 * env + err
phen2 = b0_2 + b1_2 * env + err

scenario1 = data.frame("env" = rep(env,2),
                       "gen" = rep(c("G1","G2"),each = 5),
                       "phen" = c(phen1,phen2))
plot1 = ggplot(data = scenario1, aes(x = env, y = phen, group = gen, colour = gen)) + 
  geom_line() + theme_classic() + ggtitle("Cogradient Variation, No GxE") +
  geom_segment(aes(x = 0, y = 1.0, xend = 20, yend = 5.0), colour = "black", linetype = "dashed")


# Cogradient, GxE

b0_1 = 1 #Intercept Group 1
b0_2 = 3 #Intercept Group 2
b1_1 = 0.1 #Slope Group 1
b1_2 = 0.2 #Slope Group 2

phen1 = b0_1 + b1_1 * env + err
phen2 = b0_2 + b1_2 * env + err

scenario2 = data.frame("env" = rep(env,2),
                       "gen" = rep(c("G1","G2"),each = 5),
                       "phen" = c(phen1,phen2))
plot2 = ggplot(data = scenario2, aes(x = env, y = phen, group = gen, colour = gen)) + 
  geom_line() + theme_classic() + ggtitle("Cogradient Variation, with GxE") +
    geom_segment(aes(x = 0, y = 1.0, xend = 20, yend = 7.0), colour = "black", linetype = "dashed")


# Countergradient, No GxE
b0_1 = 1 #Intercept Group 1
b0_2 = 3 #Intercept Group 2
b1_1 = 0.1 #Slope Group 1
b1_2 = 0.1 #Slope Group 2

phen1 = b0_1 + b1_1 * env + err
phen2 = b0_2 + b1_2 * env + err

scenario3 = data.frame("env" = rep(env,2),
                       "gen" = rep(c("G1","G2"),each = 5),
                       "phen" = c(phen1,phen2))
plot3 = ggplot(data = scenario3, aes(x = env, y = phen, group = gen, colour = gen)) + 
  geom_line() + theme_classic() + ggtitle("Countergradient Variation, No GxE")+
  geom_segment(aes(x = 0, y = 3.0, xend = 20, yend = 3.0), colour = "black", linetype = "dashed")

# Countergradient, GxE

b0_1 = 1 #Intercept Group 1
b0_2 = 3 #Intercept Group 2
b1_1 = 0.1 #Slope Group 1
b1_2 = 0.2 #Slope Group 2

phen1 = b0_1 + b1_1 * env + err
phen2 = b0_2 + b1_2 * env + err

scenario4 = data.frame("env" = rep(env,2),
                       "gen" = rep(c("G1","G2"),each = 5),
                       "phen" = c(phen1,phen2))
plot4 = ggplot(data = scenario4, aes(x = env, y = phen, group = gen, colour = gen)) + 
  geom_line() + theme_classic() + ggtitle("Countergradient Variation, with GxE")+
  geom_segment(aes(x = 0, y = 3.0, xend = 20, yend = 3.0), colour = "black", linetype = "dashed")

# No Gradient, Just GxE
b0_1 = 1 #Intercept Group 1
b0_2 = 5 #Intercept Group 2
b1_1 = 0.2 #Slope Group 1
b1_2 = -0.2 #Slope Group 2

phen1 = b0_1 + b1_1 * env + err
phen2 = b0_2 + b1_2 * env + err

scenario5 = data.frame("env" = rep(env,2),
                       "gen" = rep(c("G1","G2"),each = 5),
                       "phen" = c(phen1,phen2))
plot5 = ggplot(data = scenario5, aes(x = env, y = phen, group = gen, colour = gen)) + 
  geom_line() + theme_classic() + ggtitle("No Gradient, just GxE")

grid.arrange(plot1,plot3,plot2,plot4,plot5,nrow = 3)

```

```{r Original Function to create simulated data}

### This function creates reciprocal transplant data ###
### returns the data, true cov(G,E), obs cov(G,E) ###

simulateGV <- function(G_diff_N1_N2, E_diff_N1_N2, n, GxE=c(0,0,0,0)){ 
  ## G_diff_N1_N2 genetic difference in phenotype from population N1 to N2
  ## E_diff_N1_N2 environment difference in phenotype from population N1 to N2
  ## n is the sample size per population-environment combo
  
  (G <- c(-G_diff_N1_N2/2, G_diff_N1_N2/2)) 
  (E <- c(-E_diff_N1_N2/2, E_diff_N1_N2/2))
  (covGE_true <- cov(G, E))

  P1_E1 <- G[1] + E[1] + GxE[1] + rnorm(n,0,1)
  P1_E2 <- G[1] + E[2] + GxE[2] + rnorm(n,0,1)
  P2_E1 <- G[2] + E[1] + GxE[3] + rnorm(n,0,1)
  P2_E2 <- G[2] + E[2] + GxE[4] + rnorm(n,0,1)

  phen <- c(P1_E1, P1_E2, P2_E1, P2_E2)
  gen <- rep(c("G1","G2"), each=n*2)
  env <- rep(c("E1", "E2", "E1", "E2"), each=n)

  data_GE <- data.frame(phen, gen, env)

  (covGE_data <- cov(c(mean(phen[gen=="G1"]),
                       mean(phen[gen=="G2"])), 
                     c(mean(phen[env=="E1"]), 
                       mean(phen[env=="E2"]))
    ))
    
  return(list(data_GE=data_GE, 
              covGE_true=covGE_true, 
              covGE_data=covGE_data)) 
              #corGE_true = corGE_true, corGE_data=corGE_data))
}

test = simulateGV(0.5,0.5,10,GxE=c(0,0,0,0))

```

```{r Bootstrap Function for original simulated data}

## Bootstrap to estimate error

bootstrapGV <- function(data_GE){

  G_means <- NULL
  E_means <- NULL
  for (j in 1:nlevels(data_GE$gen)){
    cond_G <- data_GE$gen==levels(data_GE$gen)[j]
    G_means[j] <- mean(sample(data_GE$phen[cond_G], size=sum(cond_G), replace=TRUE))
    
    cond_E <- data_GE$env==levels(data_GE$env)[j]
    E_means[j] <- mean(sample(data_GE$phen[cond_E], size=sum(cond_E), replace=TRUE))
  }
  covGE_boot <- cov(G_means, E_means)
  #corGE_boot <- cor(G_means, E_means)
  return(list(covGE_boot))#, corGE_boot))
}

# Bootstrapped mean and confidence intervals
new_dat <- replicate(1000, bootstrapGV(data_GE), simplify=TRUE) #Repeat this same function to get permutation null dist.
(cov_avg = mean(as.numeric(unlist(new_dat))))  # should be similar to actual average 
(cov_CI = quantile(unlist(new_dat), probs=c(0.025, 0.975), type=1)) # Get confidence intervals around covariance

```



```{r Old Simulation}
setwd("~/Documents/GitHub/CnGV/src/")

#cores=detectCores()
#cl <- makeCluster(cores[1]-1) #not to overload your computer
#registerDoParallel(cl)

nrep = seq(from = 1, to = 5, by = 1)
samplesize = seq(from = 1, to = 10, by = 1) # Must be Integer
diff_effect = seq(from = 1, to = 5, by = 1) 
#row = 0 # for indexing
#nsims = 7 # Multiply against rep for total number sims

## Parallelize

#foreach(s = 1:nsims, .combine = rbind) %dopar% { 
  
## Establish datasets for output  

simout = data.frame()
power_data = data.frame()

## Simulation 

for(j in 1:length(nrep)){ 
  for(k in 1:length(samplesize)) {
    for(l in 1:length(diff_effect)) {
      
       # Set parameters 
       rep = nrep[j]
       val = samplesize[k] 
       effect_val = diff_effect[l]
      
       cat(j,k,l, "\n")  # Counter 
       
       # Run functions
       simout_temp <- simulateGV(effect_val, effect_val, val) #Run function to generate data 
       sampling_dist_cov <- replicate(1000, bootstrapGV(simout_temp$data_GE)[1], simplify=TRUE) #Bootstrap
       CI = quantile(unlist(sampling_dist_cov), probs=c(0.025, 0.975), type=1) #Get confidence intervals
       
       overunder = 0
       if (CI[1] <= 0) {overunder = 0} else 
         if (CI[2] <= 0) {overunder = 0} else 
           overunder = 1

       # Data output  
       simoutk = data.frame("replicate" = rep,
                       "samplesize" = val,
                       "GE_Diff" = effect_val,
                       "covGE_true" = simout_temp$covGE_true,
                       "covGE_data" = simout_temp$covGE_data, 
                       "lwrCI" = CI[1],
                       "uprCI" = CI[2],
                       "overunder" = overunder)
        simout = rbind(simout,simoutk)
        
    }
  }
  power_data = GetPower(simout)
 # write.csv(simout,paste("SimResults_06122019",s,".csv",sep="_"),row.names=FALSE)
}
  
## To stop:
#stopCluster(cl)
```

*Colate Data from Power analysis simulations*

```{r}
# Use below only running in parallel. 

## List all the files

temp <- list.files(pattern= "*06102019")
temp1 = do.call(rbind, lapply(temp, function(x) read.csv(x, stringsAsFactors = FALSE)))

## Add column for replicate

nrep = (length(diff_effect)*length(samplesize))
totsim = (nsims*length(rep))
temp1$replicate = rep(1:totsim, each = nrep)  

```


### Meta-Analysis Data Analysis ###

Code for Real Data: 

For population level, need function to establish G_means and E_means. 
1. This info should be in factor form. 
2. May need to translate it into quantitative form?

In all cases, need to standardize phenotype before estimation.
Standardize = (Data - avg(data))/std dev(data)

Still to do: (7/2/2019)
 - Incorporate GxE?
 - Ensure fits other datasets
 - Match to simulation code?
```{r}
## Load Data 

falc_dat = read.csv("~/Documents/GitHub/CnGV/data/Falconer_template.csv")


## Standardize and format data
standardize_fun <- function(rawdat){
  dat_avg = mean(rawdat$phen) 
  dat_std = sd(rawdat$phen)
  rawdat$phen_corrected = ((rawdat$phen-dat_avg)/dat_std)
  rawdat$env = as.factor(rawdat$env)
return(rawdat)
}

simdata = standardize_fun(simlin)
falcdata = standardize_fun()
```
 
```{r}
## Gradient level analyses

Gradient_level_fun  <- function(z){
  
  Cov_matrix = data.frame() # Covariance matrix

  # Environmental parameters
  E_hat = mean(z$exp_env) # Deal with integer problem later
  
  # Model
  grand_mod = lm(phen_corrected ~ exp_env + gen_factor, data = z) #additive vs. interaction

  # Predict phenotypic value for each genotype across <all> environments
  temp_dat <- expand.grid(gen_factor = unique(z$gen_factor),
                          exp_env = seq(from = as.integer(min(z$exp_env)), to = as.integer(max(z$exp_env),by=1)),
                          phen_corrected = 0)
  mm = model.matrix(terms(grand_mod),temp_dat)
  temp_dat$phen_corrected = mm %*% coef(grand_mod) # Predicted phenotypes

  # G_means
  G_means = filter(temp_dat, exp_env == as.integer(E_hat))
  
  # E_means
  E_mean_temp = data.frame()
  for(j in 1:length(unique(z$nat_env_mean))){
    E = as.integer(unique(z$nat_env_mean)[j])
    E_temp = temp_dat[which(temp_dat$exp_env == E),]
    E_mean = mean(E_temp$phen_corrected)
    E_mean_temp1 = data.frame("E_means" = E_mean,
                              "exp_env" = unique(E_temp$exp_env))
    E_mean_temp = rbind(E_mean_temp, E_mean_temp1)
  }
  
  # Data output
  Cov_matrix = data.frame("gen" = unique(z$gen),
                          "env" = unique(z$nat_env_mean),
                          "G_means" = G_means$phen_corrected,
                          "E_means" = E_mean_temp[,1])
  
  return(as.data.frame(Cov_matrix))
}

```


```{r}

## Generate predicted phenotypes for each G and E
covariance_data = Gradient_level_fun(sim_dat)

## Covariance from data
G <- c(covariance_data$G_means) 
E <- c(covariance_data$E_means) 
(covGE <- cov(G, E)) # For Falconer data, actual covariance = 0.51

```

```{r}
## Shuffle data for bootstrapping
raw_dat = falc_dat
shuffle_fun <- function(raw_dat){ 
    
  raw_dat2 = split(raw_dat,raw_dat$gen)
  temp_covmat = data.frame()

  for(n in 1:length(raw_dat2)){
      phen_temp = raw_dat2[[n]][sample(nrow(raw_dat2[[n]]), size = nrow(raw_dat2[[n]]),replace = TRUE),]
      temp_covmat = rbind(temp_covmat,phen_temp)
    }
  return(temp_covmat)
} 

```



```{r}
# Combine functions

gen_cov_fun <- function(dat){
  
  shuffle_dat = shuffle_fun(dat) #Shuffle
  mean_dat = Gradient_level_fun(shuffle_dat) #Get G and E means
  cov_est_temp = cov(mean_dat$G_means,mean_dat$E_means) #Estimate covariance
  
  return(cov_est_temp)
}

# Bootstrapped mean and confidence intervals
new_dat <- replicate(100, gen_cov_fun(falc_dat), simplify=TRUE) #Repeat this same function to get permutation null dist.
(cov_avg = mean(new_dat))  # should be similar to actual average 
(cov_CI = quantile(unlist(new_dat), probs=c(0.025, 0.975), type=1)) # Get confidence intervals around covariance

# Actual mean 
real_dat = Gradient_level_fun(falc_dat)
obs = (cov_est_temp = cov(real_dat$G_means,real_dat$E_means))
```

```{r}
## Permutation Function
nulldat = falc_dat

Perm_fun <- function(nulldat){
  
  nulldat2 = split(nulldat,nulldat$gen)
  temp_nullmat = data.frame()
  
  for(i in 1:length(nulldat2)){

  null_temp = sample(nulldat2[[i]][,"phen_corrected"], size = nrow(nulldat2[[i]]), replace=FALSE) 
  temp_nullmat. <- data.frame("gen_factor" = nulldat2[[i]][,"gen_factor"],
                              "env" = nulldat2[[i]][,"env"],
                              "phen_corrected" = null_temp)
  temp_nullmat = rbind(temp_nullmat, temp_nullmat.)
  }
  return(as.data.frame(temp_nullmat))
}
```

```{r}
null_cov_fun <- function(dat){
  
  shuffle_dat = Perm_fun(dat) #Shuffle
  mean_dat = Gradient_level_fun(shuffle_dat) #Get G and E means
  cov_null_temp = cov(mean_dat$G_means,mean_dat$E_means) #Estimate covariance
  
  return(cov_null_temp)
}

Perm_fun(falc_dat)

null_dist <- replicate(100, Perm_fun(falc_dat), simplify=TRUE) 
null_dist <- as.data.frame(unlist(null_dist))
hist(null_dist$)
# Permutation Hypothesis testing
(rank(c(obs,new_dat))[1])/(length(new_dat)+1) # outside distribution (p = 0.0). Greater than 0.5, 1-value. Less than, keep. Probability of observing by chance

#Plotting will add C.I.s from bootstrap

```

### Extra code ###

Keeping in case code is useful in future. 

```{r Extra Code, eval = FALSE}
## Was a part of code used in for-loop above, but presently unused. 
rownames(sampling_dist_cov) <- c("covGE_boot", "corGE_boot")
sampling_dist_cov <- data.frame(as.matrix(unlist(t(sampling_dist_cov))))

# cor
quantile(unlist(sampling_dist_cov[2,]), probs=c(0.025, 0.975), type=1)
# this is always 1 or -1 because we only have 2 data points

boxplot(phen~gen)

var(phen)

```

Below is Molly's initial attempt at sims - ended up not taking this approach in favor of the above linear model approach.

```{r ExtraCode2, eval = FALSE}
G = scenario2$G
E = scenario2$E
n = scenario2$n
GxE = scenario2$GxE

simulateGV <- function(G, E, n, GxE){ 
  ## G are genetic differences in phenotype 
  ## E is environment difference in phenotype between populations
  ## n is the sample size per population-environment combo
  
  # Calculate true covariance
  (covGE_true <- cov(G, E))

  new_data = data.frame()
  # Simulate new data
  for(i in 1:length(G)){
    for(j in 1:length(E)){
      for(k in 1:length(GxE)){
        
    simdat <- G[i] + E[j] + GxE[k] + rnorm(n,0,1) #Is this correct - sd of 1?
    new_data. <- data.frame("gen" = G[i],
                            "gen_factor" = rep(paste("G",order(G)[i],sep="_"), each = nrow(simdat)),
                            "env" = E[j],
                            "env_factor" = rep(paste("E",order(E)[j],sep="_"), each = nrow(simdat)),
                            "GxE" = GxE[k],
                            "phen" = simdat)
    new_data = rbind(new_data, new_data.)
      }
    }
  }

  data_GE <- data.frame("phen" = new_data$phen, 
                        "gen" = new_data$gen_factor, 
                        "env" = new_data$env_factor)
#generalize
  for(l in 1: )
  (covGE_data <- cov(c(mean(phen[gen=="G_1"]),
                       mean(phen[gen=="G_2"])), 
                     c(mean(phen[env=="E_1"]), 
                       mean(phen[env=="E2"]))
    ))
    
  return(list(data_GE=data_GE, 
              covGE_true=covGE_true, 
              covGE_data=covGE_data)) 
              #corGE_true = corGE_true, corGE_data=corGE_data))
}

ggplot(data = data_GE, aes(y=phen,x=env,group = gen)) + geom_smooth() + theme_classic()
```